---
title: "covKCD" 
author: "Peter Hoff"
date: "`r format(Sys.time(), '%Y-%m-%d')`"  
output: github_document 
--- 


### The Kronecker-core decomposition 

Let $Y_1,\ldots,Y_n$ be a sample of $p_1\times p_2$ matrices, and 
let $y_1,\ldots,y_n$ be their vectorizations. The sample covariance 
matrix $S$ of $y_1,\ldots, y_n$ is a $p\times p$ positive semidefinite 
matrix, where $p=p_1 \cdot p_2$. 
The Kronecker-core decomposition $(K,C)$ of the covariance matrix $S$ is 
a matrix factorization of the form 
\[
S = (K_2\otimes K_1)^{1/2} C (K_2\otimes K_1)^{1/2}, 
\]
where  

* $K_1$ is a $p_1\times p_1$ matrix representing 
across-row covariance of the data matrices;  
* $K_2$ is a $p_2\times p_2$ matrix representing 
across-column covariance of the data matrices; 
* $C$ is a core covariance matrix with zero
average across-row and across-column covariance. 

The covariance matrix $K=K_2\otimes K_1$ represents
the "separable part" of $S$, whereas $C$ represents the "nonseparable part". 
In particular, if $S$ is separable, than $C=I_p$. 

### Reference

Hoff, McCormack and Zhang (2022). [Core Shrinkage Covariance Estimation for Matrix-variate Data.](https://arxiv.org/abs/2207.12484)


### Installation

```{r,eval=FALSE} 
# Development version 
devtools::install_github("pdhoff/covKCD")   

# CRAN-approved version 
install.packages("covKCD") 
```


### Basic usage

```{r}
p1<-3 ; p2<-4 ; n<-10 

# Random sample of n independent standard normal p1xp2 matrices 
Y<-array(rnorm(n*p1*p2),dim=c(n,p1,p2)) 
dim(Y)


# Sample covariance matrix 
S<-covKCD::mcov(Y) 
dim(S) 


# KCD 
KCDS<-covKCD::covKCD(S,p1,p2) 

KCDS$K1 

KCDS$K2


# Check decomposition 
K1h<-covKCD::msqrt(KCDS$K1) 
K2h<-covKCD::msqrt(KCDS$K2) 

range( kronecker(K2h,K1h) %*% KCDS$C %*% kronecker(K2h,K1h) - S )
```

### Additional examples

* [Simulation study from Hoff, McCormack and Zhang (2022).](https://www2.stat.duke.edu/~pdh10/Code/arXiv.2207.12484/)

* [Speech recognition example from Hoff, McCormack and Zhang (2022).](https://www2.stat.duke.edu/~pdh10/Code/arXiv.2207.12484/)
  



